---
title: "Analyses of New Workers in a Firm"
author: "A.Mayuri(2348133)"
date: "2023-09-05"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

# Problem Statement:
Take a real life data having variables as continuous and categorical. Consider it as population and draw a sample of appropriate size by applying the SRS technique and give the estimates for Population mean and population total for both continuous and categorical variables. Write a report on it. 
Report should include the following:
Complete data description with proper definition of study characteristics. Formula of each estimates with procedure and Analysis using R programming. Complete interpretation. significance of study in real life

## Data Description:

We have a dataset of size 206 (Large sample) where it contains the information about the employees who have joined the MNC firm branch in India in the last 6 months.

## Variable of Interest: 

we have cgpa,age,hours of work,salary,expenditure are the quantitative data. while gender and additional exams taken are qualitative data

## Objective:

1) Understand the data holistically - Using summary function
2) use Descriptive analyses to express the data -Graphical representation
3) estimate the suitable sample size for the study
4) estimate population mean and population total for the available      quantitative variables available.
5) Categorical Data analyses

# ANALYSIS

## Step1: 
let us initialize and attach the dataset to the given R-Markdown.

### Code

```{r }
library(readxl)
Book <- read_excel("C:/Users/mayur/Desktop/Mstat/tri sem 1/R/Book.xlsx")
View(Book)
attach(Book)

```

## Step2: 
Descriptive data analyses: let us try to understand the given variables.

note: the summary gives a "character" as a remark when the variable is categorical in nature.

### Code

```{r }
summary(Book)


```
#### Interpretation

thus from the summary we have found the population mean for different continuous variables and the distribution of the data within different quantiles.
Mean age= 23.85 
Mean hours of work per day   : 6.689 
Mean CGPA  : 7.597 
Mean salary   : 9.68
Mean expenditure  :5.99

# Step2: Graph
we can use different graphs to pictorially represent the data

### Code
```{r }
#Pie chart for gender
m=table(Book$GENDER)
pie(m,main='Gender distribution', col=c('pink','blue'))

#Bar chart for hours of work
n=table(Book$`Hrs of work per day`)
barplot(n, main="Hours of work per week",col=c("lavender","purple","green","yellow"))

# Histogram chart for salary information
p=table(Book$`salary in lakhs`)
hist(p,main="Salary in lakhs",col=c("blue","red","yellow","pink","purple","black","white","orange","lavender"))


```
## Step3:
We have to calculate the sample size for the study 

### Code
```{r }
population=Book$`Hrs of work per day`
N=length(Book$`Hrs of work per day`)
N

p=sd(population)
p
S=N/(N-1)*(p^2)
S
sqrt(3.4936) #we have S as variance or S^2 in formula thus we tk sqrt of it to find d=z(a/2)*SE
d=1.96*1.869117
d
n=((N*(S^2))*3.84)/((3.84*S^2)+(N*d^2))
n

```
#### Interpretation:
Thus we need an approximate minimum of sample size of  35.60221~36

## Step4:

##### Hours of Work per day
Here we are trying to find if the sample mean is an unbiased estimator of the population, evaluvate the population total for the quantitaive data "hours of work per day"

### Code
```{r }
pop_mean=mean(Book$`Hrs of work per day`)
pop_mean

n=36
set.seed(12)
sampDist1 <- replicate(20000,mean(sample(Book$`Hrs of work per day`, n)))
mean(sampDist1)

Sd_srswor=sd(sampDist1)*((N-n/(N*n))^(1/2))
Sd_srswor

Pop_total=mean(sampDist1)*N
Pop_total

```
#### Interpretation: 
Thus the sample mean is an unbiased estimator of the population mean where,
sample mean=6.688961
population mean= 6.68932
ie, on an average the employees work about 6~7 hours a day
and population total (Y) = 1377.926 is the total hours worked by all the employees worked together in a day.

##### CGPA

Here we are trying to find if the sample mean is an unbiased estimator of the population, evaluvate the population total for the quantitative data "CGPA"

### Code
```{r }
pop_mean2=mean(Book$CGPA)
pop_mean2

n=36
set.seed(12)
sampDist2 <- replicate(20000,mean(sample(Book$CGPA, n)))
mean(sampDist2)

Sd_srswor2=sd(sampDist2)*((N-n/(N*n))^(1/2))
Sd_srswor2

Pop_total2=mean(sampDist2)*N
Pop_total2

```
Thus the sample mean is an unbiased estimator of the population mean where,
sample mean=7.597087
population mean= 7.597117
ie, on an average the employee GPA is about 7.6.
with a standard deviation of about 3.090771


##### Salary in Lakhs

Here we are trying to find if the sample mean is an unbiased estimator of the population, evaluvate the population total for the quantitative data "Salary in Lakhs"

### Code
```{r }
pop_mean3=mean(Book$`salary in lakhs`)
pop_mean3

n=36
set.seed(12)
sampDist3 <- replicate(20000,mean(sample(Book$`salary in lakhs`, n)))
mean(sampDist3)

Sd_srswor3=sd(sampDist3)*((N-n/(N*n))^(1/2))
Sd_srswor3

Pop_total3=mean(sampDist3)*N
Pop_total3

```

Thus the sample mean is an unbiased estimator of the population mean where,
sample mean=9.679612
population mean=  9.680476
ie, on an average the employee GPA is about 9.7 Lakhs or (9~10)Lakhs.
with a standard deviation of about 5.253572
Or the population total is about 1994.178 Lakhs which could also be interpreted as the expenditure of the firm on human capital

##### Expenditure

Here we are trying to find if the sample mean is an unbiased estimator of the population, evaluvate the population total for the quantitative data "Expenditure of th eemployee in a year"

### Code
```{r }
pop_mean4=mean(Book$`expenditure per year in lakhs`)
pop_mean4

n=36
set.seed(12)
sampDist4 <- replicate(20000,mean(sample(Book$`expenditure per year in lakhs`, n)))
mean(sampDist4)

Sd_srswor4=sd(sampDist4)*((N-n/(N*n))^(1/2))
Sd_srswor4

Pop_total4=mean(sampDist4)*N
Pop_total4

```
sample mean= 5.989806
population mean=  5.994517
ie, on an average the employee GPA is about 5.9 Lakhs or ~(6)Lakhs.
with a standard deviation of about  5.293196
or the population total is about 1234.871 lakhs per annum. or the employees spend about 1234.871 Lakhs per year.


## Categorical Data Analyses

```{r}
library(dplyr)
df1= Book %>% group_by(GENDER,`wrote any additional exams`) %>%
  summarise(mean_age=mean(age),mean_Hrs_work=mean(`Hrs of work per day`),mean_cgpa =mean(CGPA), mean_Salary=mean(`salary in lakhs`), mean_expenditure=mean(`expenditure per year in lakhs`),.groups='drop') %>%
  as.data.frame()
df1
```

Conclusion: 
From the above table we can see the matrices that is generated by grouping the means of different quantitative paramenter with respect to each qualitative option(yes/no or F/M)
ie, for eg,
Exam vs No Additional Exam
we have the mean age as to be 23.84 for females who did not take up any additional exams vs, the mean age of females who wrote exams as 24.11. This can be understood where students stay a little longer in education when they take up additional exams. also we can see that their average salary is considerably more (10.17 lakhs > 8.43 Lakhs) per annum for those who take up additional exams tend to earn more even in their first job.

Female Vs Male
We significantly observe that a female is earning less despite taking up additional exams such that of a male conterpart. we observe that the difference may not be significant in the lower strata however as they climb the ladder of a firm the difference could be significant (male-12.14 Lakhs (vs) Female=10.17 Lakhs)


#### Significance of Study:

It is a true upholding that biased employment opportunities, salary etc are towards women, differently abled and similarly weaker section exsist. it is vital to understand and take measures to counterfeit it.
Statistics as a science and tool for research helps us to validate the exsisting patterns, which can be addressed if collective measures are take.

